Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-04-19 Manuscript/manuscript.Rmd
manuscript edits

We then investigated which genomic annotations were most predictive of TAD boundary formation, across resolution for random forest models with the best given set of data-level characteristics determined above (SMOTE re-sampled with distance type predictors). We found that specific transcription factor binding sites, including SMC3ab, ZNF143, CTCF, and RAD21, were given consistently high rankings of predictive importance in the models built on the GM12878 cell line (Table 1). These are known factors associated with the hypothesized loop extrusion model [@fudenberg2016formation; @alipour2012self] and were found to be ranked consistently high on the K562 cell line as well, with similar rankings across resolutions (Supplementary Table 6). Other annotations known to be associated with gene regulation and affect structural properties of chromatin, such as the YY1  transcription factor, as well DNAse hypersensitive sites, regions of heterochromatin, and the H3k27 trimethylation histone modifications were ranked high on the 10 kb resolution, but were less consistently ranked at lower resolutions for both cell lines. <!--Moreover, the ranking profiles of genomic annotations for both cell lines were found to be more similar between the 10 kb and 25 kb resolutions, as well as between the 50 kb and 100 kb resoluions (this is proved when looking at a clustered heatmap).-->

>>>>

We then investigated which genomic annotations were most predictive of TAD boundary formation, across resolution for random forest models with the best given set of data-level characteristics determined above (SMOTE re-sampled with distance type predictors). We found that specific transcription factor binding sites (TFBS), including SMC3ab, ZNF143, CTCF, and RAD21, were given consistently high rankings of predictive importance in the models built on the GM12878 cell line (Table 1). These are known factors associated with the hypothesized loop extrusion model [@fudenberg2016formation; @alipour2012self] and were found to be ranked consistently high on the K562 cell line as well, with similar rankings across resolutions (Supplementary Table 6). Other annotations known to be associated with gene regulation and affect structural properties of chromatin, such as the YY1  transcription factor, as well DNAse hypersensitive sites, regions of heterochromatin, and the H3k27 trimethylation histone modifications were ranked high on the 10 kb resolution, but were less consistently ranked at lower resolutions for both cell lines. <!--Moreover, the ranking profiles of genomic annotations for both cell lines were found to be more similar between the 10 kb and 25 kb resolutions, as well as between the 50 kb and 100 kb resoluions (this is proved when looking at a clustered heatmap).-->

Given the prevalence of TFBS as the most consistently top ranked predictive features across resolution and cell line, we ventured to further compare performances of annotation specific models. After applying SMOTE re-sampling with distance type predictors, random forest models built on sets of histone modifications, chromatin state segmentations, and TFBS were compared to a model that included all three types together. For the GM12878 cell line, models built simply on TFBS were found to perform just as well as models that included all 3 sets of annotations according to F1-Score and MCC, consistently across resolution (Figure 6). Similar results were found when evaluating the K562 cell line (supplementary Figure 11).
